---
title: "Customer Clustering"
author: "Utkarsh Sharma"
date: "16/11/2019"
output: 
  prettydoc::html_pretty:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Problem Statement, Dataset and Pre-processing
To handle the task of clustering customers, we are taking into account a fictitious database. Customers of the business are measured in the following criteria:

1. Demographics (Gender, Age, Annual Income)
2. Channels (Acquisition Source)
3. Engagement (Spending Score)

The Spending Score is a score on a scale of 1-100 which the business uses to rank its customers. A higher spending score means the customer engages more with the products of the business, and drives more profit than a customer with a lower spending score.

```{r}
# Importing the dataset
library(readxl)
dataset = read_excel("Customers.xlsx")
dataset = dataset[2:6]

head(dataset)
```

```{r}
# Encoding the categorical variables as factors
dataset$`Acquisition Source` = as.numeric(factor(dataset$`Acquisition Source`,
                                      levels = c('Internet', 'Agent'),
                                      labels = c(1, 2)))

dataset$Gender = as.numeric(factor(dataset$Gender,
                                   levels = c('Female', 'Male'),
                                   labels = c(1, 2)))
```

## 2. K-Means Clustering
To find the optimal number of disjoint clusters to divide the customers into, we use the Silhouette method.

```{r warning = FALSE, message = FALSE}
# Optimal number of clusters
library(factoextra)
fviz_nbclust(dataset, kmeans, method='silhouette')
```

```{r}
set.seed(100)
kmeans = kmeans(x = dataset, centers = 6)
dataset$Cluster <- as.factor(kmeans$cluster)
```

## 3. Analysing Clusters

```{r warning = FALSE, message = FALSE}
library(dplyr)
cluster_analysis <- dataset %>% group_by(Cluster) %>% summarise(Count = n(),
                                                                MeanAnnualIncome = mean(`Annual Income (k$)`),
                                                                AverageAge = mean(Age),
                                                                AverageSpendingScore = mean(`Spending Score (1-100)`))

head(cluster_analysis)
```

```{r}
# Visual Analysis
library(ggplot2)
ggplot(cluster_analysis, aes(Cluster, AverageSpendingScore)) + geom_col()
```

It can be seen that clusters 4 and 5 are the clusters with the highest mean Spending Score, followed by Cluster 2. To look into the demographics of these clusters further;

```{r}
ggplot(dataset, aes(`Annual Income (k$)`, `Spending Score (1-100)`, color = Cluster)) + geom_point()
ggplot(dataset, aes(Age, `Spending Score (1-100)`, color = Cluster)) + geom_point()
```

The clusters can be summarised as;

  1. Cluster 1 - Customers with age greater than 30 and average annual income have an average Spending Score of about 50. 
  2. Cluster 2 - Young customers with low annual income have a higher than average Spending Score of about 63.
  3. Cluster 3 - Customers with age less than 60 and a higher than average annual income have a very low average Spending Score.
  4. Cluster 4 - People in their 30's with very high annual income have a very high mean Spending Score.
  5. Cluster 5 - People in their 30's with higher than average annual income have the highest mean Spending Score.
  6. Cluster 6 - People past their 40's with very low annual income have a very low average Spending Score.

## 4. Results and Future Work
The customers are successfully divided into six disjoint groups, analysis of which can help the business drive strategy to target a particular customer group to increase profits. Furthermore, we can use Classification algorithms to assign future customers into relevant groups.







